---
title: "Visualization"
subtitle: "Part I"
author: JingYao Geng
date: '`r format(Sys.time(), "%Y-%m-%d") `'
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df

```


## Basic scatterplot
Create my first scatterplot ever.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

New approach, same plot.
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```


Save and edit a plot object
```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()
```


## Adcanced scatterplot

Start with the same one and make it fancy
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```

What about the 'aes' placement...?
```{r}
weather_df %>%
ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) + #not good
  geom_smooth()
```

Let's facet some things!
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.2, size = 0.3) + # 20% transparency of the points, size = 0.3 makes the points size smaller
  geom_smooth(se = FALSE, size = 2) + # here the 'size' is to adjust the line
  facet_grid(. ~ name) # multi-panel plot. In this case, put a '.' means nothing defines the row,and put a tilde and tell it which variable defines the column
```

Let's combine some elements and try a new plot
```{r}
weather_df %>%
ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

## Some small notes

How many geoms have to exist?

You can have whatever geomes you want.

```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
```

You can use a neat geom!
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex() # a 2d histogram, where brighter colors say there's a lot of data points. Overkill for dataset that's got a thousand obs. This sort of plot are helpful for understanding what the data distribution might look like.
  # or 
  # geom_bin2d() same as geom_hex but with squares
  # or
  # geom_density2d() it's sort of like a contour plot
```

## Univariate plots
Histograms are really great.
```{r}
weather_df %>%
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Can we add color....
```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
  
# or 
#weather_df %>%
  #ggplot(aes(x = tmin, fill = name)) +
  #geom_histogram()  
  #facet_grid(., ~name)

 # or don't need the color
#weather_df %>%
  #ggplot(aes(x = tmin)) +
  #geom_histogram()  
  #facet_grid(., ~name)
```


Let's try a new geometry!
```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_density()  

weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.3)  

```

What about box plots??
```{r}
weather_df %>%
  ggplot(aes(y = tmin)) +
  geom_boxplot()

weather_df %>%
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
  
```

Trendy plots :)
```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = 0.5)
  stat_summary(fun = "median")
```

Ridge plot -- the most popular plot of 2017
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = name)) +
  geom_boxplot()
```

## Save and Embed

Let's save a scatterplot
```{r}
weather_plot = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5)

##ggsave("./results/weather_plot.pdf", weather_plot, width = 8, height = 5)
```

What about embedding...
```{r fig.width=6, fig.height=3}
weather_plot #adjust all the figures dimension
```

embed at different size
```{r fig.width=12, fig.height=6}
weather_plot
```

