---
title: "Visulization"
Subtitle: "Part II"
author: JingYao Geng
output: github_document
---

```{r setup}
ls()
library(tidyverse)
library(patchwork)
```

## load the weather data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Remeber this plot .. ?
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
  
```


## labels
title, x, y labels, and caption
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
  labs(
    title = " Temperature plot", 
    x = "Minimum daily temperature(C)", 
    y = "Maximum daily temperature (C)", 
    caption = "Data from rnoaa package; temperature in 2017."
  )
```

## Scales

Start with the same plot, x and y scales
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
  labs(
    title = " Temperature plot", 
    x = "Minimum daily temperature(C)", 
    y = "Maximum daily temperature (C)", 
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
    
  # define some scales:  
  scale_x_continuous(
    breaks = c(-15, 0, 15),  # x_axis tick marks: -15, 0, 15
    labels = c("-15 C", "0", "15")
  ) + 
  # if your x var. is categorical/factor var., you might use scale_x_discrete()
  scale_y_continuous(
    trans = "sqrt" # square-root transformation to y-axis
    # trans = "log"  log-transformation
    # position = "right" put the y-labels on the right-side
  )
```

Look at color scales

I have defined a mapping b/w name variable and the color aesthetic but I'm not supper happy with this, there are a couple of things that we can do to update this.

scale_color is gonna be another one of these that controls the color attributes, it controls the color feature, and there are gonna to be a lot of options.

 - scale_color_hue(h = c(0, 360): describe a feature of colors, h tells u sth about the hue of a scale, and by default from 0 to 360.
 - In scale_color_hue and scale color other things, u can rename the sort of in scale x
 - scale_color_viridis() assme that u are putting in is a continuous value, if it's discrete var. u need to specify: discrete = TRUE
 - google: viridis color palettes, for various color blindness, this is more accessible

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
  labs(
    title = " Temperature plot", 
    x = "Minimum daily temperature(C)", 
    y = "Maximum daily temperature (C)", 
    caption = "Data from rnoaa package; temperature in 2017."
  ) + 
    viridis::scale_color_viridis(
      name = "Location",
      discrete = TRUE
      
    )
      
      #scale_color_hue(
      #name = "Location",
      #h = c(100,300))
    
```

Next, move away from things that control specific mappings or find grain control of the mappings b/w var. and the plot features.
Plot features that aren't data dependent that just like background color and font choices. -- Themes:

## Themes
Shift the legend to the bottom.
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
  labs(
    title = " Temperature plot", 
    x = "Minimum daily temperature(C)", 
    y = "Maximum daily temperature (C)", 
    caption = "Data from rnoaa package; temperature in 2017."
  ) + 
    viridis::scale_color_viridis(
      name = "Location",
      discrete = TRUE
    ) +
    theme(legend.position = "bottom")
  
```

Change the overall theme
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
  labs(
    title = " Temperature plot", 
    x = "Minimum daily temperature(C)", 
    y = "Maximum daily temperature (C)", 
    caption = "Data from rnoaa package; temperature in 2017."
  ) + 
    viridis::scale_color_viridis(
      name = "Location",
      discrete = TRUE
      ) +
    #theme_bw() # background of the graph
    theme_minimal() + # same as the theme_bw, without the box arounded
    # theme_classic()
    # ggthemes::theme_economist()
    # ggthemes::theme_excel() There are a lot of themes in ggthemes
    theme(legend.position = "bottom")
```

## Setting options
use in the very beginning of the document
- theme_set() to determine whatever theme I wanna exist everywhere inside of this document
- options() are to override ggplots default coloring.  viridis palette
```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk(
  fig.width = 6,
  fig.asp = 0.6, # aspect ratio: the height over the width
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continous.color = "viridis", 
  ggplot2.continous.fill = "viridis"
) 

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data args in 'geom'

```{r}
central_park = 
  weather_df %>%
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>%
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tamx, color = name)) +
  geom_point() +
  #geom_line()
  geom_line(data = cental_park)
```

## 'patchwork'

remeber faceting?

```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~name)
```

What happen when you want multipanel plots but can't facet...?

The individual plots here matter less than the idea that potentially what I would wanna be able to do is organize all of these into a single figure, and that's where patchwork comes in is if for example I wanted to put a scatter plot 'tmax_tmin_p" next to a density polt "prcp_dens_p", or have that be organized around another plot "tmax_date_p" that has sort og fundamentally different arguments and geometries and mappings. 
```{r}
tmax_tmin_p = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "none")

prcp_dens_p =
  weather_df %>%
  filter(prcp > 0) %>%
  ggplot(aes(x = prcp, fill = name )) +
  geom_density(apha = 0.5) +
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>%
  ggplot(aes(x =  date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

#tmax_tmin_p + prcp_dens_p # a two side by side panels
#tmax_tmin_p + prcp_dens_p + tmax_date_p # 3 panels side by side
#tmax_tmin_p + (prcp_dens_p + tmax_date_p) # 3 panels side by side, but more space for the first
#tmax_tmin_p / (prcp_dens_p + tmax_date_p) # first on the top, the other 2 panels below
(tmax_tmin_p + prcp_dens_p) / tmax_date_p # two panels on the top, one on the bottom
```

## Data manipulation

Control your factors
```{r}
weather_df %>%
  mutate(
    name = factor(name), 
    name = forcats::fct_relevel(name, c("Wakiki_HA")) # put wakiki on the first
  ) %>%
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

What if I wanted densities for tmin and tmax simultaneously?
 - could I get on a single panel, could I take one panel and have the tmin and then have the tmax

```{r}
weather_df %>%
  filter(name == "CentralPark_NY") %>%
  pivot_longer(
    tmax:tmin, 
    names_to = "observation", 
    values_to = "temperature"
  ) %>%
  ggplot(aes(x = temperature, fill = observation)) +
  geom_density(alpha = 0.5) # a temperature density plot for tmin and tmax
```

```{r}
weather_df %>%
  pivot_longer(
    tmax:tmin, 
    names_to = "observation", 
    values_to = "temperature"
  ) %>%
  ggplot(aes(x = temperature, fill = observation)) +
  geom_density(alpha = 0.5) + 
  facet_grid(. ~name)
```


Additional example: 
1:08:30, 
I skipped it.


